Packaging research artefacts with RO-Crate

### 2022-01-01 latex/ro-crate.tex
escape {}

\begin{listing}
%\includegraphics{210053i02}
\footnotesize
\begin{verbatim}
{ "@context": "https://w3id.org/ro/crate/1.1/context",
  "@graph": [
    { "@id": "ro-crate-metadata.json",      
      "@type": "CreativeWork",
      "conformsTo": {"@id": "https://w3id.org/ro/crate/1.1"},
      "about": {"@id": "./"}
    },
    { "@id": "./",
      "@type": "Dataset",
      "name": "A simplified RO-Crate",
      "author": {"@id": "#alice"},
      "license": {"@id": "https://spdx.org/licenses/CC-BY-4.0"},
      "datePublished": "2021-11-02T16:04:43Z",
      "hasPart": [
        {"@id": "survey-responses-2019.csv"},
        {"@id": "https://example.com/pics/5707039334816454031_o.jpg"}
      ]
    },
    { "@id": "survey-responses-2019.csv",
      "@type": "File",
      "about": {"@id": "https://example.com/pics/5707039334816454031_o.jpg"},
      "author": {"@id": "#alice"}
    },
    { "@id": "https://example.com/pics/5707039334816454031_o.jpg",
      "@type": ["File", "ImageObject"],
      "contentLocation": {"@id": "http://sws.geonames.org/8152662/"},
      "author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
    },
    { "@id": "#alice",
      "@type": "Person",
      "name": "Alice"
    },
    { "@id": "https://orcid.org/0000-0002-1825-0097",
      "@type": "Person",
      "name": "Josiah Carberry"
    },
    { "@id": "http://sws.geonames.org/8152662/",
      "@type": "Place",
      "name": "Catalina Park"
    },
    { "@id": "https://spdx.org/licenses/CC-BY-4.0",
      "@type": "CreativeWork",
      "name": "Creative Commons Attribution 4.0"
    }
  ]
}    
\end{verbatim}    
\caption{%\qq{}\protect\footnotemark[31]
Simplified\protect\footnote{Recommended properties for types shown in
Listing \ref{lis1} also include
\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
\texttt{description},
\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
\texttt{keywords},
\texttt{publisher}; these properties and corresponding contextual
entities are
excluded here for brevity. See complete example
\url
{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.}
RO-Crate metadata file showing the
flattened compacted JSON-LD \texttt{@graph} array containing the data entities
and contextual entities, cross-referenced using \texttt{@id}. The
\texttt{ro-crate-metadata.json} entity self-declares conformance with the
RO-Crate specification using a versioned persistent identifier, further
RO-Crate descriptions are on the root data entity \texttt{./} or any of the
referenced data or contextual entities, as exemplified by the data
entity \texttt{ImageObject} referencing contextual entities for
\texttt{contentLocation} and \texttt{author} that differs from that of the overall
RO-Crate. In this crate, \texttt{about} of the CSV data entity reference the
\texttt{ImageObject}, which then take the roles of both a data entity and
contextual entity. While \texttt{Person} entities ideally are identified with
ORCID PIDs as for Josiah, \texttt{\#alice} is here in contrast an RO-Crate
local identifier, highlighting the pragmatic ``just enough'' Linked
Data approach.}\label{lis1}
\end{listing}

In\footnotetext{Recommended properties for types shown in
Listing \ref{lis1} also include
\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
\texttt{description},
\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
\texttt{keywords},
\texttt{publisher}; these properties and corresponding contextual
entities are
excluded here for brevity. See complete example
\url
{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{{entity}} are directly
under \texttt{@graph} and represents the RDF triples with a common \textit{subject}
(\texttt{@id}), mapped \textit{properties} like \texttt{hasPart}, and \textit{objects} -- as
either literal \texttt{"string"} values, referenced \texttt{{objects}} (which
properties are listed in its own entity), or a JSON \texttt{[list]} of these.
If processed as JSON-LD, this forms an RDF graph by matching the \texttt{@id}
IRIs and applying the \texttt{@context} mapping to Schema.org terms.
\normalsize

@@

Similarly, while the Schema.org context
currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
have \texttt{"@type": "@id"} annotations for object properties, RO-Crate
JSON-LD distinguishes explicitly between references to other entities
(\texttt{{"@id": "\#alice"}})
and string values (\texttt{"Alice"}) -- meaning
RO-Crate applications can find references for corresponding entities
and IRIs without parsing the \texttt{@context} to understand a particular
property. Notably this is exploited by the \texttt{ro-crate-html-js}
\cite{ro-crate-html-js} tool to provide reliable HTML rendering for
otherwise unknown properties and types.

>>>>

\begin{listing}
%\includegraphics{210053i02}
\footnotesize
\begin{verbatim}
{ "@context": "https://w3id.org/ro/crate/1.1/context",
  "@graph": [
    { "@id": "ro-crate-metadata.json",      
      "@type": "CreativeWork",
      "conformsTo": {"@id": "https://w3id.org/ro/crate/1.1"},
      "about": {"@id": "./"}
    },
    { "@id": "./",
      "@type": "Dataset",
      "name": "A simplified RO-Crate",
      "author": {"@id": "#alice"},
      "license": {"@id": "https://spdx.org/licenses/CC-BY-4.0"},
      "datePublished": "2021-11-02T16:04:43Z",
      "hasPart": [
        {"@id": "survey-responses-2019.csv"},
        {"@id": "https://example.com/pics/5707039334816454031_o.jpg"}
      ]
    },
    { "@id": "survey-responses-2019.csv",
      "@type": "File",
      "about": {"@id": "https://example.com/pics/5707039334816454031_o.jpg"},
      "author": {"@id": "#alice"}
    },
    { "@id": "https://example.com/pics/5707039334816454031_o.jpg",
      "@type": ["File", "ImageObject"],
      "contentLocation": {"@id": "http://sws.geonames.org/8152662/"},
      "author": {"@id": "https://orcid.org/0000-0002-1825-0097"}
    },
    { "@id": "#alice",
      "@type": "Person",
      "name": "Alice"
    },
    { "@id": "https://orcid.org/0000-0002-1825-0097",
      "@type": "Person",
      "name": "Josiah Carberry"
    },
    { "@id": "http://sws.geonames.org/8152662/",
      "@type": "Place",
      "name": "Catalina Park"
    },
    { "@id": "https://spdx.org/licenses/CC-BY-4.0",
      "@type": "CreativeWork",
      "name": "Creative Commons Attribution 4.0"
    }
  ]
}    
\end{verbatim}    
\caption{%\qq{}\protect\footnotemark[31]
Simplified\protect\footnote{Recommended properties for types shown in
Listing \ref{lis1} also include
\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
\texttt{description},
\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
\texttt{keywords},
\texttt{publisher}; these properties and corresponding contextual
entities are
excluded here for brevity. See complete example
\url
{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.}
RO-Crate metadata file showing the
flattened compacted JSON-LD \texttt{@graph} array containing the data entities
and contextual entities, cross-referenced using \texttt{@id}. The
\texttt{ro-crate-metadata.json} entity self-declares conformance with the
RO-Crate specification using a versioned persistent identifier, further
RO-Crate descriptions are on the root data entity \texttt{./} or any of the
referenced data or contextual entities. This is exemplified by the data
entity \texttt{ImageObject} referencing contextual entities for
\texttt{contentLocation} and \texttt{author} that differs from that of the overall
RO-Crate. In this crate, \texttt{about} of the CSV data entity reference the
\texttt{ImageObject}, which then take the roles of both a data entity and
contextual entity. While \texttt{Person} entities ideally are identified with
ORCID PIDs as for Josiah, \texttt{\#alice} is here in contrast an RO-Crate
local identifier, highlighting the pragmatic ``just enough'' Linked
Data approach.}\label{lis1}
\end{listing}

In\footnotetext{Recommended properties for types shown in
Listing \ref{lis1} also include
\texttt{affiliation}, \texttt{citation}, \texttt{contactPoint},
\texttt{description},
\texttt{encodingFormat}, \texttt{funder}, \texttt{geo}, \texttt{identifier},
\texttt{keywords},
\texttt{publisher}; these properties and corresponding contextual
entities are
excluded here for brevity. See complete example
\url
{https://www.researchobject.org/2021-packaging-research-artefacts-with-ro-crate/listing1/}.} this flattened profile of JSON-LD, each \texttt{\{entity\}} is directly
under \texttt{@graph} and represents the RDF triples with a common \textit{subject}
(\texttt{@id}), mapped \textit{properties} like \texttt{hasPart}, and \textit{objects} -- as
either literal \texttt{"string"} values, referenced \texttt{\{objects\}} (which
properties are listed in its own entity), or a JSON \texttt{[list]} of these.
If processed as JSON-LD, this forms an RDF graph by matching the \texttt{@id}
IRIs and applying the \texttt{@context} mapping to Schema.org terms.
\normalsize

@@

Similarly, while the Schema.org context
currently\footnote{\url{https://schema.org/version/13.0/schemaorg-current-http.jsonld}}
have \texttt{"@type": "@id"} annotations for object properties, RO-Crate
JSON-LD distinguishes explicitly between references to other entities
(\texttt{\{"@id": "\#alice"}\})
and string values (\texttt{"Alice"}) -- meaning
RO-Crate applications can find references for corresponding entities
and IRIs without parsing the \texttt{@context} to understand a particular
property. Notably this is exploited by the \texttt{ro-crate-html-js}
\cite{ro-crate-html-js} tool to provide reliable HTML rendering for
otherwise unknown properties and types.
